	\documentclass[a4paper,12pt]{article}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{eurosym}
\usepackage{vmargin}
\usepackage{amsmath}
\usepackage{graphics}
\usepackage{framed}
\usepackage{epsfig}
\usepackage{subfigure}
\usepackage{enumerate}
\usepackage{fancyhdr}

\setcounter{MaxMatrixCols}{10}
%TCIDATA{OutputFilter=LATEX.DLL}
%TCIDATA{Version=5.00.0.2570}
%TCIDATA{<META NAME="SaveForMode"CONTENT="1">}
%TCIDATA{LastRevised=Wednesday, February 23, 201113:24:34}
%TCIDATA{<META NAME="GraphicsSave" CONTENT="32">}
%TCIDATA{Language=American English}

\pagestyle{fancy}
\setmarginsrb{20mm}{0mm}{20mm}{25mm}{12mm}{11mm}{0mm}{11mm}
\lhead{MS4222} \rhead{Kevin O'Brien} \chead{Sampling} %\input{tcilatex}

\begin{document}
%=========================================%

\section*{Sampling With and Without Replacement}
\subsection*{Sampling without replacement}
		\begin{itemize}
			\item Sampling is said to be ``without replacement" when a unit is selected at random from the population and it is not returned to the main lot. \item The first unit is selected out of a population of size $N$ and the second unit is selected out of the remaining population of  $N-1$ units and so on.
			\item For example, if you draw one card out of a deck of 52, there are only 51 cards left to draw from if you are selecting a second card.
		\end{itemize}
		

\subsection*{Example}
A lot of 100 semiconductor chips contains 20 that are defective.
Two chips are selected at random, without replacement from the lot.
\begin{itemize}
	\item What is the probability that the first one is defective? \\(Answer: 20/100, i.e 0.20)
	\item What is the probability that the second one is defective given that the first one was defective? \\(Answer: 19/99)
	\item What is the probability that the second one is defective given that the first one was not defective? \\(Answer: 20/99)
\end{itemize}
	
		

\subsection*{Sampling With Replacement }

Sampling is called ``with replacement" when a unit selected at random from the population is returned to the population and then a second element is selected at random. Whenever a unit is selected, the population contains all the same units.
\begin{itemize}
	\item What is the probability of guessing a PIN number for an ATM card at the first attempt.
	
	\item Importantly a digit can be used twice, or more, in PIN codes.
	
	\item For example $1337$ is a valid pin number, where $3$ appears twice.
	
	\item
	We have a one-in-ten chance of picking the first digit correctly, a one-in-ten chance of the guessing the second, and so on.
	
	\item All of these events are independent, so the probability of guess the correct PIN is $0.1 \times 0.1 \times 0.1 \times 0.1 = 0.0001$
\end{itemize}	




\end{document}		
