\documentclass[a4paper,12pt]{article}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{eurosym}
\usepackage{vmargin}
\usepackage{amsmath}
\usepackage{graphics}
\usepackage{epsfig}
\usepackage{subfigure}
\usepackage{framed}
\usepackage{enumerate}
\usepackage{fancyhdr}

\setcounter{MaxMatrixCols}{10}
%TCIDATA{OutputFilter=LATEX.DLL}
%TCIDATA{Version=5.00.0.2570}
%TCIDATA{<META NAME="SaveForMode"CONTENT="1">}
%TCIDATA{LastRevised=Wednesday, February 23, 201113:24:34}
%TCIDATA{<META NAME="GraphicsSave" CONTENT="32">}
%TCIDATA{Language=American English}

\pagestyle{fancy}
\setmarginsrb{20mm}{0mm}{20mm}{25mm}{12mm}{11mm}{0mm}{11mm}
\lhead{MS4222} \rhead{Kevin O'Brien} \chead{Normal Distribution} %\input{tcilatex}

\begin{document}

%---------------------------------------------------------------------%

\section*{Poisson Example}
A motor dealership which specializes in agricultural machinery sells one vehicle every 2 days, on average. Using Poisson Distribution Tables, answer the following questions.
\begin{enumerate}[(a)]
\item   What is the probability that the dealership sells at least one vehicle in one particular day?
\item   What is the probability that the dealership will sell exactly one vehicle in one particular day?
\item   What is the probability that the dealership will sell 4 vehicles or more in a six day working week?
\end{enumerate}


%---------------------------------------------------------------------%

\noindent \textbf{Solutions}\\
Expected Occurrences per Day: m = 0.5
\begin{enumerate}[(a)]
 
\item  Probability that the dealership sells at least one vehicle in one particular day?
\[ P(X \geq 1) = 0.3935 \]
\item  Probability that the dealership will sell exactly one vehicle in one particular day?
\[ P(X = 1) = P(X \geq 1) - P(X \geq 2)  = 0.3935 - 0.0902 = 0.3031\]
\item  Probability that the dealership will sell 4 vehicles or more in a six day working week?
\begin{itemize}
\item  For a 6 day week, m=3
\item  $P(X \geq 4) = 0.3528$
\end{itemize}
\end{enumerate}


\end{document}
