\documentclass[a4paper,12pt]{article}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{eurosym}
\usepackage{vmargin}
\usepackage{amsmath}
\usepackage{graphics}
\usepackage{epsfig}
\usepackage{framed}
\usepackage{subfigure}
\usepackage{enumerate}
\usepackage{fancyhdr}

\setcounter{MaxMatrixCols}{10}
%TCIDATA{OutputFilter=LATEX.DLL}
%TCIDATA{Version=5.00.0.2570}
%TCIDATA{<META NAME="SaveForMode"CONTENT="1">}
%TCIDATA{LastRevised=Wednesday, February 23, 201113:24:34}
%TCIDATA{<META NAME="GraphicsSave" CONTENT="32">}
%TCIDATA{Language=American English}

\pagestyle{fancy}
\setmarginsrb{20mm}{0mm}{20mm}{25mm}{12mm}{11mm}{0mm}{11mm}
\lhead{MS4222} \rhead{Kevin O'Brien} \chead{Discrete Probability Distributions} %\input{tcilatex}

\begin{document}


%---------------------------------------------------------------------%

\section*{The Cumulative Distribution Function}
\begin{itemize}
\item  The Cumulative Distribution Function, denoted $F(x)$, is a common way that the probabilities
of a random variable (both discrete and continuous) can be summarized.
\item  The Cumulative Distribution Function, which also can be
described by a formula or summarized in a table, is defined as:
\[F(x) = P(X \leq x) \]
\item  The notation for a cumulative distribution function, F(x), entails using a capital
"F".  
\begin{itemize}
\item[$\bullet$] The notation for a probability mass or density function, $f(x)$, i.e. using a lowercase "f". The notation is \textbf{not} interchangeable.
\end{itemize}
\end{itemize}

\section*{Useful Results for Working with Tables}
%(Demonstration on the blackboard re: partitioning of the sample space, using examples on next slide)
\begin{itemize}
\item  $P(X \leq 1) = P(X=0) + P(X=1)$
\item  $P(X \leq r) = P(X=0)+ P(X=1) + \ldots P(X= r)$
\item  $P(X \leq 0) = P(X=0)$
\item  $P(X = r) = P(X \geq r ) - P(X \geq r + 1)$
\item  \textbf{Complement Rule}: $P(X \leq r-1) = P(X < r) = 1 - P(X \geq r)$
\item  \textbf{Interval Rule}: $P(a \leq X \leq  b)= P(X \geq a) - P(X \geq b + 1).$
\end{itemize}
For the binomial distribution, if the probability of success is greater than 0.5, instead of
considering the number of successes, to use the table we consider
the number of failures.
\section*{Binomial Example 1}
Suppose a signal of 100 bits is transmitted and the probability of
sending a bit correctly is 0.9. What is the probability of
\begin{enumerate}[(a)]
\item  at least 10 errors
\item  exactly 7 errors
\item  Between 5 and 15 errors (inclusively).
\end{enumerate}

\subsection*{Solution}
\begin{itemize}
\item  Since the probability of success is 0.9. We consider the distribution
of the number of failures (errors).
\item  We reverse the definition of `success' and `failure'. Success is now defined as an error.
\item  The probability that a bit is sent incorrectly is 0.1.
\item  Let X be the total number of errors: $X \sim B(100, 0.1)$.
\item  From Binomal Tables
\begin{enumerate}[(a)]
    \item 
$P(X \geq 10) = 0.5487$.
\item  $P(X = 7)=P(X \geq 7) - P(X \geq 8) =0.8828 - 0.7939 = 0.0889$.
\item  $P(5 \leq X  \leq 15) = P(X \geq 5) - P(X \geq 16) =0.9763 - 0.0399 = 0.9364$
\end{enumerate}
\end{itemize}


\end{document}
