
\documentclass[]{article}

\voffset=-1.5cm
\oddsidemargin=0.0cm
\textwidth = 480pt

\usepackage{framed}
\usepackage{subfiles}
\usepackage{graphics}
\usepackage{newlfont}
\usepackage{eurosym}
\usepackage{amsmath,amsthm,amsfonts}
\usepackage{amsmath}
\usepackage{enumerate}
\usepackage{color}
\usepackage{multicol}
\usepackage{amssymb}
\usepackage{multicol}
\usepackage[dvipsnames]{xcolor}
\usepackage{graphicx}
\begin{document}
\title{MS4222}



\section*{MS4222 Tutorial Sheet 1 (Week 3)}
\begin{enumerate}
\item Answer the following short questions.
\begin{enumerate}[(a)]
\item A die is rolled and a coin is tossed, find the probability that the die shows an odd number and the coin shows a head.
\item A card is drawn at random from a deck of cards. Find the probability of getting the 3 of diamond.
\item A card is drawn at random from a deck of cards. Find the probability of getting a queen.

\item A die is rolled, find the probability that the number obtained is greater than 4. 
\item Two coins are tossed, find the probability that one head only is obtained. 

\item  Two dice are rolled, find the probability that the sum is equal to 5. 

\item  A card is drawn at random from a deck of cards. Find the probability of getting the King of heart. 
\end{enumerate}



% Answers to above exercises: 
% a) 2 / 6 = 1 / 3 
% b) 2 / 4 = 1 / 2 
% c) 4 / 36 = 1 / 9 
% d) 1 / 52



\item 
A fair coin is tossed three times. Write out the sample space $S$.

Write the sample points of the following events:

\begin{enumerate}[(a)]
\item the event that the first toss is heads,
\item the event that the second toss is heads,
\item exactly two heads in a row.
\end{enumerate}
    \item 
A couple has two children; the sample space is $S = {bb, bg, gb, gg}$ with probability 0.25 for each sample point. Find the probability $p$ that both children are boys if it is known that: 
\begin{enumerate}[(a)]
\item at least one of the children is a boy,
\item the older child is a boy.
\end{enumerate}
    \item 
Compute the probability $P$ of each event:
\begin{enumerate}[(a)]
\item An even number appears in the toss of a fair die;
\item  One or more tails appear in the toss of three fair coins;
\item  A blue marble appears in a random drawing of one marble from a box containing four white, three blue, and five red marbles.
\end{enumerate}
    \item Answer the following short questions
\begin{enumerate}[(a)]
\item Suppose a student is selected at random from 100 students where 30 are taking mathematics, 20 are taking computer science, and 10 are taking mathematics and computer science. Find the probability $p$ that the student is taking mathematics or computer science.
\item 
The blood groups of 200 people is distributed as follows: 50 have type A blood, 65 have B blood type, 70 have O blood type and 15 have type AB blood. If a person from this group is selected at random, what is the probability that this person has O blood type?
\end{enumerate}

    \item 
A lot contains 13 items of which 4 are defective. Three items are drawn at random from the lot one after the other. Find the probability $p$ that all three are nondefective.
    \item 
Suppose we are given the following information about events A and B.
\begin{itemize}
    \item $P(A) = 0.45$
    \item $P(B) = 0.35$
    \item $P(A \cup B) = 0.65$
    
\end{itemize}
\begin{enumerate}[(a)]
\item Compute $P(A \cap B)$.
\end{enumerate}
\item     
The probability that A hits a target is 1/3 and the probability that B hits a target is 1/5.  They both fire at the target. Find the probability that:

\begin{enumerate}[(a)]
\item  A does not hit the target; 
\item  both hit the target; 
\item  one of them hits the target; 
\item  neither hits the target.
\end{enumerate}

    \item 
A machine is composed of 2 components A and B, which function independently of one another. The machine works only if both components are working. The probability that component A works is 0.98 and the probability that component B works is 0.95.

\begin{enumerate}[(a)]
\item	What is the probability that the machine works, 
\item	What is the probability that both components have failed.
\end{enumerate} 



\item Two dice are rolled, find the probability that the sum is 
\begin{enumerate}[(a)]
\item equal to 1 

\item equal to 4 

\item less than 13
\end{enumerate}




    
    
    
    \item Components of a certain type are shipped to a supplier in batches of ten. Suppose that 52\% of all such batches contain no defective components, 31\% contain one defective component, and 17\% contain two defective components. Two components from a batch are randomly selected and tested. What are the probabilities associated with 0, 1, and 2 defective components being in the batch under each of the following conditions? (Round your answers to four decimal places.)
    
    
\item 
Six horses are entered in a race. Larry picks two of the horses at random, and bets on them. 
How many ways are there of picking two horses at random from six?

\item 

Suppose a fair coin is tossed six times. The number of heads which can occur with their respective
probabilities follows:
\[
\begin{tabular}{|c|c|c|c|c|c|c|c|} \hline
$X_i$	 & 0	& 1	& 2	& 3	& 4	& 5	& 6 \\ \hline
$P(X_i)$ &	1/64 &	6/64 &	15/64 &	20/64 &	15/64 &	6/64 &	1/64 \\ \hline
\end{tabular}
\]

\begin{enumerate}[(a)]
\item	Compute the expected value (i.e. expected number of heads).
\item	Compute the variance of the heads.
\end{enumerate}
%---------------------------------------------------%

\item A player tosses two fair coins. He wins \$2 if two heads occur, and \$1 if one head occurs. On the other hand, he loses \$3 if no heads occur. Find the expected value $E(X)$ of the game, where $X$ is the winnings.
\smallskip
\textit{(The game is fair, favorable, or unfavorable to the player according as E(X) = 0,$E(X) > 0$ or $E(X) < 0$.)}

    \item 
On completion of a programming project, four programmers from a team submit a collection of subroutines to an acceptance group. The following table shows the percentage of subroutines each programmer submitted and the probability that a subroutine submitted by each programmer will pass the certification test based on historical data.
\[
\begin{tabular}{|c|c|c|c|c|} \hline
Programmer &	A	& B	& C&	D \\ \hline \hline
Proportion of & 0.10 &	0.20 &	0.30 &	0.40 \\ 
subroutines submitted &&&&	 \\ \hline
Probability &	0.55 &	0.60 &	0.95 &	0.75 \\ 
of acceptance &&&&\\ \hline
\end{tabular}
\]
\begin{enumerate}[(a)]
\item
What is the proportion of subroutines that pass the acceptance test?
\item After the acceptance tests are completed, one of the subroutines is selected at random and found to have passed the test. What is the probability that it was written by Programmer A?
\end{enumerate}
    \item 
Two manufacturing plants produce similar parts. Plant A produces 1000 parts, 100 of which are defective. Plant B produces 2000 parts, 170 of which are defective. 

\begin{enumerate}[(a)]
\item	Present this information using a contingency table
\item	A part is selected at random and found to be defective. What is the probability that it came from plant A ?
\end{enumerate}

\item The probability distribution of a discrete random variable is be tabulated as follows

\begin{center}
\begin{tabular}{|c||c|c|c|}
\hline
$x_i$  & 1 & 2 & 3  \\\hline
$p(x_i)$ & 0.4 & 0.3 & 0.3 \\ \hline

\end{tabular}
\end{center}

\begin{enumerate}[(a)]
\item What is expected value $E(X)$?
\end{enumerate}


\item The probability distribution of a discrete random variable is be tabulated as follows

\begin{center}
\begin{tabular}{|c||c|c|c|c|c|c|}
\hline
$x_i$  & 1 & 2 & 3 & 4 & 5 & 6 \\\hline
$p(x_i)$ & 2/8 & 1/8& 1/8 & 1/8& c & 1/8\\
\hline
\end{tabular}
\end{center}

\begin{enumerate}[(a)]
\item What is the value of $c$?
\item What is expected value and variance of the outcomes?
\end{enumerate}


\item 

A lottery with 500 tickets gives one prize of \$100, three prizes of \$50 each, and five prizes of \$25 each.
\begin{enumerate}[(a)]
\item  Find the expected winnings of a ticket. 
\item If a ticket costs \$1, what is the expected value of the game?
\end{enumerate}
\item 
A random sample with replacement of size n = 2 is drawn from the set $\{1, 2, 3\}$, yielding the following
9-element sample space:

\[S = {(1, 1), (1, 2), (1, 3), (2, 1), (2, 2), (2, 3), (3, 1), (3, 2), (3, 3)}\]

Each pairing is equally probable. Let $X$ denote the sum of the two numbers. 

\begin{enumerate}[(a)]
\item Construct a table that contains all the possible values of X and the probability of that outcome.

\item Compute the expected value E(X).
\end{enumerate}


\item 

Suppose X is a random variable with mean 15 and standard deviation 5.

\begin{enumerate}[(a)]
\item Compute the expected value of $X^2$ 
\end{enumerate}
\textit{(Recall: standard deviation is the square root of variance.)}
    
    \item 
For a particular Java assembler interface, the operand stack size has the
following probabilities:

\begin{center}
\begin{tabular}{|c||c|c|c|c|c|}
\hline
% after \\: \hline or \cline{col1-col2} \cline{col3-col4} ...
Stack Size  & 0 & 1 & 2 & 3 & 4 \\ \hline
Probability & 0.15 & 0.05 & 0.10 &0.20 &0.50\\
\hline
\end{tabular}
\end{center}


\begin{enumerate}[(a)]

\item Calculate the expected stack size.
\item Calculate the variance of the stack size.

\end{enumerate}

\item 
The random variables X and Y have the following joint distribution
\begin{center}
\begin{tabular}{|c||c|c|c|} \hline
	& Y=0	& Y=1 & 	 Y=2 \\ \hline \hline
X=0	& 2/10	& 1/10	& 2/10 \\ \hline
X=1	& 0/10	& 2/10	& 1/10 \\ \hline
X=2	& 1/10	& 0/10	& 1/10 \\ \hline
 
\end{tabular} 
\end{center}
 

\begin{enumerate}[(a)]
\item	Calculate the marginal distributions of X and Y.     
\item	What is the expected value of X?
\item	What is the expected value of Y?                 
\end{enumerate} 



\item 
Suppose X has the following probability mass function: $P(0) =
0.2$, $P(1) = 0.5$, $P(2) = 0.3$. Calculate $E[X]$ and $E[X^2]$.

\newpage

\item Consider the random variables $X$ and $Y$. $X$ takes the values 0,1 and 2. $Y$ takes the values $0$ and $1$.
The joint probabilities for each pair are given by the following table.
\begin{center}
\begin{tabular}{|c|c|c|c|}
\hline  & $X=0$ & $X=1$ & $X=2$  \\ 
\hline $Y=0$ & 0.1  & 0.4 & 0.1 \\ 
\hline  $Y=1$ & 0.1 & 0.1 & 0.2 \\ 
\hline 
\end{tabular} 
\end{center}
\begin{enumerate}[(a)]
\item Compute the expected values of $X$ and $Y$.
\item Compute the $E(X|Y=1)$
\end{enumerate}

\item Consider the random variables $X$ and $Y$. Both $X$ and $Y$ take the values 0,$\;$1 and 2. 
The joint probabilities for each pair are given by the following table.
\begin{center}
\begin{tabular}{|c|c|c|c|}
\hline  & $X=0$ & $X=1$ & $X=2$ \\ 
\hline $Y=0$ & 0.1 & 0.15 & 0.1 \\ 
\hline $Y=1$ & 0.1 & 0.1 & 0.1 \\ 
\hline $Y=2$ & 0.2 & 0.05 & 0.1 \\ 
\hline 
\end{tabular} 
\end{center}
Compute the $E(U)$ expected value of $U$, where $U=X-Y$.
\end{enumerate}
\end{document}


%%-- https://www.examsolutions.net/tutorials/exam-questions-discrete-random-variables/
