	\documentclass[a4paper,12pt]{article}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{eurosym}
\usepackage{vmargin}
\usepackage{amsmath}
\usepackage{graphics}
\usepackage{epsfig}
\usepackage{subfigure}
\usepackage{enumerate}
\usepackage{fancyhdr}

\setcounter{MaxMatrixCols}{10}
%TCIDATA{OutputFilter=LATEX.DLL}
%TCIDATA{Version=5.00.0.2570}
%TCIDATA{<META NAME="SaveForMode"CONTENT="1">}
%TCIDATA{LastRevised=Wednesday, February 23, 201113:24:34}
%TCIDATA{<META NAME="GraphicsSave" CONTENT="32">}
%TCIDATA{Language=American English}

\pagestyle{fancy}
\setmarginsrb{20mm}{0mm}{20mm}{25mm}{12mm}{11mm}{0mm}{11mm}
\lhead{MS4222} \rhead{Kevin O'Brien} \chead{Continuous Probability Distributions} %\input{tcilatex}

\begin{document}
\section*{Exponential Distribution: Review Questions}
\begin{enumerate}
\item
An average of five calls per hour are received by a machine repair department. 
Beginning the observation at any point in time, what is the probability that the first call for service will arrive within five minutes?

\item Jobs are sent to a printer at an average of 5 jobs per hour. 
\begin{enumerate}[(a)]
	\item What is the expected time between jobs? 
	\item What is the probability that the next job is sent within 6 minutes after the previous job?
\end{enumerate}

\item 
Assume that the time, denominated in minutes, between arrivals of customers at a particular bank is exponentially distributed with a rate parameter of 0.25. 

\begin{enumerate}[(a)]
    \item What is the mean duration between arrivals?
	\item
	Find the probability that the time between arrivals is greater than 5 minutes.
	\item]
	Find the probability that the time between arrivals will be less than 2 minute.
\end{enumerate}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\item Suppose that customers arrive at a filling station at the rate of 3 per hour. Given that a customer has just arrived, the time is takes for the next customer to arrive is called a \textbf{waiting time}. Let $T$ be the symbol for the ``waiting times" variable. 


Suppose that a customer has just arrived.
\begin{enumerate}[(a)] \item What is the expected waiting time between customer arrivals? 
\subitem Compute $E(T)$
\item What is the variance of waiting times? \subitem Compute $\mbox{Var}(T)$.
	\item What is the probability that the next customer will arrive within the next fifteen minutes?
	\item What is the probability that no customers arrive in the next half hour?
\end{enumerate}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%




\end{enumerate}

\end{document}
